Quantum Information

### 2024-12-03 examples/paper.tex
Some definitions

@@

@@

\begin{proposition}
  Let $A\in\mathcal{M}_d$ be any square matrix, then 
  \begin{equation*}
     (A \otimes \id) \ket{\Omega} = (\id\otimes A^T) \ket{\Omega}.
  \end{equation*}
\end{proposition}
\begin{proof}
   Exercise: explicit calculation.
\end{proof}
More generally, let $d$ and $D$ be natural numbers, $\ket{\Omega_d}$ and $\ket{\Omega_D}$ be the $d$- and the $D$ dimensional maximally entagled states. Then
\begin{proposition}
  Let $A\in\mathcal{M}_{d\times D}$ be a $d\times D$ matrix, then 
  \begin{equation*}
    (A \otimes \id) \ket{\Omega_D} = (\id\otimes A^T) \ket{\Omega_d}.
  \end{equation*}
\end{proposition}
\begin{proof}
  Exercise: explicit calculation.
\end{proof}
A special case of this equation is if $d=1$. Then $\ket{\Omega_d} \in \mathbb{C}^d\otimes \mathbb{C}^d \simeq \mathbb{C}$ is just $1$, and the $d\times D$ matrix is just a linear functional. Writing out this case explicitly, we obtain 
\begin{proposition}
  Let $\ket{\Psi}\in\mathbb{C}^{d}$ and $\ket{\Omega}\in \mathbb{C}^d \otimes \mathbb{C}^d$ be the maximally entangled state. Then 
  \begin{equation*}
    (\bra{\Psi} \otimes \id) \ket{\Omega} = \ket{\bar{\Psi}},
  \end{equation*}
  the complex conjugate of $\ket{\Psi}$.
\end{proposition}
\begin{proof}
  Exercise: explicit calculation.
\end{proof}

@@

At the beginning of the protocol, Alice and Bob have the state $\ket{\Psi}\otimes\ket{\Omega}\in \mathcal{H}_{A'} \otimes \mathcal{H}_{A}\otimes \mathcal{H}_B$; in Alice's lab the qubits $A'$ and $A$, in Bob's lab the qubit $B$. The protocol consists of the following steps:
\begin{enumerate}
  \item Alice measures the qubits $A$ and $A'$ in the Bell basis, and obtains outcome $(\alpha,\beta)\in\{0,1\}^2$.
  \item  Alice sends the classical bits $\alpha$ and $\beta$ to Bob
  \item  Bob applies on the particle $B$ the operator $x^\alpha Z^\beta$. 
\end{enumerate}
We will show that at the end of the protocol, Bob obtains the state $\ket{\Psi}$. For that, note the post-measurement state after Alice's measurement, up to normalization, is 
\begin{equation*}
   \ket{\Phi_{\alpha\beta}} = (\ket{\Omega_{\alpha\beta}}\bra{\Omega_{\alpha\beta}} \otimes\id) (\ket{\Psi} \otimes \ket{\Omega}),
\end{equation*}
where the order of the tensor components is $A'AB$. As the measurement is rank-1, the outcome is of the form $\ket{\Phi_{\alpha\beta}} = \ket{\Omega_{\alpha\beta}}\otimes \ket{\Psi_{\alpha\beta}}$, where
\begin{equation*}
     \ket{\Psi_{\alpha\beta}} = (\bra{\Omega_{\alpha\beta}} \otimes\id) (\ket{\Psi} \otimes \ket{\Omega}).
\end{equation*}
 Let us write out now the definition of $\ket{\Omega_{\alpha \beta}}$ in this formula:
\begin{equation*}
  \ket{\Psi_{\alpha\beta}} = (\bra{\Omega_{\alpha\beta}} \otimes\id) (\ket{\Psi} \otimes \ket{\Omega}) = (\bra{\Omega} \otimes\id) (\id \otimes Z^\beta X^\alpha\otimes\id) (\ket{\Psi} \otimes \ket{\Omega}).
\end{equation*}
Using now {\color{red} ... Property} of the maximally entangled state, we obtain that 
\begin{equation*}
  \ket{\Psi_{\alpha\beta}} =  (\bra{\Omega} \otimes\id) (\id \otimes Z^\beta X^\alpha \otimes\id) (\ket{\Psi} \otimes \ket{\Omega}) =  (\bra{\Omega} \otimes\id) (\id \otimes \id \otimes X^\alpha Z^\beta ) (\ket{\Psi} \otimes \ket{\Omega}) = X^\alpha Z^\beta\ket{\Psi_{00}},
\end{equation*}
where 
\begin{equation*}
  \ket{\Psi_{00}} =  (\bra{\Omega} \otimes\id) (\ket{\Psi} \otimes \ket{\Omega}).
\end{equation*}
Using again {\color{red} ... Property} of the entangled state, we obtain that
\begin{equation*}
  \ket{\Psi_{00}} =  (\bra{\Omega} \otimes\id) (\ket{\Psi} \otimes \ket{\Omega}) = (\bra{\bar{\Psi}} \otimes\id) \ket{\Omega} = \ket{\Psi}.
\end{equation*}
We thus obtain that the state after Alice's measurement, if she obtains ouctcome $\alpha\beta$, is
\begin{equation*}
  \ket{\Phi_{\alpha\beta}} = \ket{\Omega_{\alpha\beta}}\otimes X^\alpha Z^\beta \ket{\Psi}.
\end{equation*}
Therefore, if Bob applies $Z^\beta X^\alpha$ on his part of the state, he will obtain the state $\ket{\Psi}$ independent of Alice's side.

>>>>

\section{Introduction}

\subsection{Bra-ket notation}

\begin{itemize}
  \item Bra and ket
  \item Scalar product
  \item Adjoint
  \item Eigenvalues
\end{itemize}

\subsection{Positivity}

\begin{definition}[Positivity]
  A matrix $A\in \bounded{\mathcal{H}}$ is \emph{positive semidefinite} if $\bra{\Psi} A\ket{\Psi} \geq 0$ for all $\ket{\Psi}\in\mathcal{H}$. It is \emph{positive definite} if $\bra{\Psi} A\ket{\Psi} > 0$ for all $\ket{\Psi}\in\mathcal{H}$. 
\end{definition}

\begin{theorem}
  Let $\mathcal{H}$ be a Hilbert space, $A\in\bounded{\mathcal{H}}$ be a matrix. The following are equivalent.
  \begin{itemize}
    \item $A$ is positive semidefinite (positive definite)
    \item $A$ is Hermitian and it has only non-negative (positive) eigenvalues
    \item There is a(n invertible) matrix $X\in \bounded{\mathcal{H}}$ such that $A = X^\dagger X$.
  \end{itemize}
\end{theorem}

\subsection{Tensor product}

\section{Quantum mechanics}

\subsection{The rules of quantum mechanics}
 \begin{itemize}
   \item States
   \item Time evolution
   \item Measurements: destructive, then projective
 \end{itemize}

\subsection{The rules of quantum mechanics: mixed states}
\begin{itemize}
  \item Mixed states
  \item POVMs
  \item CP maps
\end{itemize}

\begin{definition}[Ensemble]
  Let $\mathcal{H}$ be a Hilbert space. En \emph{ensemble} is a collection of normalized vectors with probabilities: it is given by $n\in\mathbb{N}$ and $\big((p_i, \ket{\Psi_i})\big)_{i=1}^n \subseteq (\mathbb{R}_{\geq 0} \times \mathcal{H})^n$ such that $\|\Psi_i\| = 1$ for all $i=1,\dots,n$ and $\sum_{i=1}^n p_i = 1$.
\end{definition}
\subsubsection{Mixed states}

\begin{definition}[Density matrix]
  Let $\mathcal{H}$ be a Hilbert space. A \emph{state} or a \emph{density matrix} on $\mathcal{H}$ is a matrix $\rho\in\bounded{\mathcal{H}}$ such that $\tr\rho = 1$, and $\rho\geq 0$, i.e., it is positive semidefinite. The state $\rho$ is called \emph{pure} if it is rank-one. 
\end{definition}

\begin{proposition}
  Let $\mathcal{H}$ be a Hilbert space and $\rho\in\bounded{\mathcal{H}}$ be a pure state. Then there is $\ket{\Psi}\in\mathcal{H}$, $\|\Psi\| = 1$, such that $\rho  = \ket{\Psi} \bra{\Psi}$.
\end{proposition}

\subsubsection{CP maps}

\begin{definition}
    Let $\mathcal{H}$ and $\mathcal{K}$ be two Hilbert spaces. Let $T:\bounded{\mathcal{H}}\to\bounded{\mathcal{K}}$ be linear. We say that $T$ is \emph{trace preserving} if $\tr T(\rho) = \tr \rho$, for all $\rho\in\bounded{\mathcal{H}}$.
\end{definition}

\begin{definition}
  Let $\mathcal{H}$ and $\mathcal{K}$ be two Hilbert spaces. Let $T:\bounded{\mathcal{H}}\to\bounded{\mathcal{K}}$ be linear. We say that $T$ is a Kraus map if there are $A_i\in \mathcal{B}(\mathcal{H}\rightarrow \mathcal{K})$ such that 
  \begin{equation}\label{eq:Kraus_decomp_def}
    T(\rho) = \sum_{i=1}^n A_i \rho A_i^\dagger.
  \end{equation}
  The operators $A_i$ are called the Kraus operators, and \cref{eq:Kraus_decomp_def} the Kraus representation of $T$.
\end{definition}

\begin{proposition}
    Let $\mathcal{H}$ and $\mathcal{K}$ be two Hilbert spaces. Let $T:\bounded{\mathcal{H}}\to\bounded{\mathcal{K}}$ be a Kraus map with Kraus operators $\{A_i\}_{i=1}^n$. Then $T$ is trace preserving iff $\sum_i A_i^\dagger A_i = \id_{\mathcal{H}}$.
\end{proposition}
\begin{proof}
  Exercise.
\end{proof}

\begin{definition}
  Let $\mathcal{H}$ and $\mathcal{K}$ be two Hilbert spaces. Let $T:\bounded{\mathcal{H}}\to\bounded{\mathcal{K}}$ be linear. We say that 
  \begin{itemize}
    \item $T$ is \emph{positivity preserving}, or simply \emph{positive}, if $T(\rho)\geq 0$ for all $\rho\in\bounded{\mathcal{H}}$. 
    \item $T$ is $n$-positive if $T\otimes\id: \bounded{\mathcal{H}}\otimes\bounded{\mathbb{C}^n} \to \bounded{\mathcal{H}} \otimes \bounded{\mathbb{C}^n}$ is positive.
    \item $T$ is \emph{completely positive (CP)} if it is $n$-positive for all $n\in\mathbb{N}$.
  \end{itemize}
\end{definition}

@@

\begin{definition}[Entanglement of pure states]
  A pure state $\ket{\Psi}\in \mathcal{H}\otimes\mathcal{K}$ is called entangled if it is not a product.
\end{definition}

\begin{definition}[Entanglement]
  A state $\rho \in \bounded{\mathcal{H}} \otimes \bounded{\mathcal{K}}$ is called \emph{separable} if there are density matrices $\mu_i\in \bounded{\mathcal{H}}$, $\nu_i\in\bounded{\mathcal{H}}$, $i=1\dots n$, and a probability distribution $p\in\mathbb{R}^n$ such that
  \begin{equation*}
    \rho = \sum_{i=1}^n p_i \cdot \mu_i \otimes \nu_i.    
  \end{equation*}
  The state $\rho$ is called \emph{entangled} if it is not separable.
\end{definition}

@@

\begin{proposition}\label{prop:max_ent_pull_through}
  Let $A\in\mathcal{M}_d$ be any square matrix, then 
  \begin{equation*}
     (A \otimes \id) \ket{\Omega} = (\id\otimes A^T) \ket{\Omega}.
  \end{equation*}
\end{proposition}
\begin{proof}
   Exercise: explicit calculation.
\end{proof}
More generally, let $d$ and $D$ be natural numbers, $\ket{\Omega_d}$ and $\ket{\Omega_D}$ be the $d$- and the $D$ dimensional maximally entagled states. Then
\begin{proposition}
  Let $A\in\mathcal{M}_{d\times D}$ be a $d\times D$ matrix, then 
  \begin{equation*}
    (A \otimes \id) \ket{\Omega_D} = (\id\otimes A^T) \ket{\Omega_d}.
  \end{equation*}
\end{proposition}
\begin{proof}
  Exercise: explicit calculation.
\end{proof}
A special case of this equation is if $d=1$. Then $\ket{\Omega_d} \in \mathbb{C}^d\otimes \mathbb{C}^d \simeq \mathbb{C}$ is just $1$, and the $d\times D$ matrix is just a linear functional. Writing out this case explicitly, we obtain 
\begin{proposition}\label{prop:max_ent_pull_through_lin_fcn}
  Let $\ket{\Psi}\in\mathbb{C}^{d}$ and $\ket{\Omega}\in \mathbb{C}^d \otimes \mathbb{C}^d$ be the maximally entangled state. Then 
  \begin{equation*}
    (\bra{\Psi} \otimes \id) \ket{\Omega} = \ket{\bar{\Psi}},
  \end{equation*}
  the complex conjugate of $\ket{\Psi}$.
\end{proposition}
\begin{proof}
  Exercise: explicit calculation.
\end{proof}

@@

At the beginning of the protocol, Alice and Bob have the state $\ket{\Psi}\otimes\ket{\Omega}\in \mathcal{H}_{A'} \otimes \mathcal{H}_{A}\otimes \mathcal{H}_B$; in Alice's lab the qubits $A'$ and $A$, in Bob's lab the qubit $B$. The protocol consists of the following steps:
\begin{enumerate}
  \item Alice measures the qubits $A$ and $A'$ in the Bell basis, and obtains outcome $(\alpha,\beta)\in\{0,1\}^2$.
  \item  Alice sends the classical bits $\alpha$ and $\beta$ to Bob
  \item  Bob applies on the particle $B$ the operator $x^\alpha Z^\beta$. 
\end{enumerate}
We will show that at the end of the protocol, Bob obtains the state $\ket{\Psi}$. For that, note the post-measurement state after Alice's measurement, up to normalization, is 
\begin{equation*}
   \ket{\Phi_{\alpha\beta}} = (\ket{\Omega_{\alpha\beta}}\bra{\Omega_{\alpha\beta}} \otimes\id) (\ket{\Psi} \otimes \ket{\Omega}),
\end{equation*}
where the order of the tensor components is $A'AB$. As the measurement is rank-1, the outcome is of the form $\ket{\Phi_{\alpha\beta}} = \ket{\Omega_{\alpha\beta}}\otimes \ket{\Psi_{\alpha\beta}}$, where
\begin{equation*}
     \ket{\Psi_{\alpha\beta}} = (\bra{\Omega_{\alpha\beta}} \otimes\id) (\ket{\Psi} \otimes \ket{\Omega}).
\end{equation*}
 Let us write out now the definition of $\ket{\Omega_{\alpha \beta}}$ in this formula:
\begin{equation*}
  \ket{\Psi_{\alpha\beta}} = (\bra{\Omega_{\alpha\beta}} \otimes\id) (\ket{\Psi} \otimes \ket{\Omega}) = (\bra{\Omega} \otimes\id) (\id \otimes Z^\beta X^\alpha\otimes\id) (\ket{\Psi} \otimes \ket{\Omega}).
\end{equation*}
Using now \cref{prop:max_ent_pull_through}, i.e., that the maximally entangled state satisfies $(A\otimes \id) \cdot \ket{\Omega} = (\id\otimes A^T) \cdot\ket{\Omega}$, for any square matrix $A$, we obtain that 
\begin{equation*}
  \ket{\Psi_{\alpha\beta}} =  (\bra{\Omega} \otimes\id) (\id \otimes Z^\beta X^\alpha \otimes\id) (\ket{\Psi} \otimes \ket{\Omega}) =  (\bra{\Omega} \otimes\id) (\id \otimes \id \otimes X^\alpha Z^\beta ) (\ket{\Psi} \otimes \ket{\Omega}) = X^\alpha Z^\beta\ket{\Psi_{00}},
\end{equation*}
where 
\begin{equation*}
  \ket{\Psi_{00}} =  (\bra{\Omega} \otimes\id) (\ket{\Psi} \otimes \ket{\Omega}).
\end{equation*}
Using now \cref{prop:max_ent_pull_through_lin_fcn} twice, we obtain that
\begin{equation*}
  \ket{\Psi_{00}} =  (\bra{\Omega} \otimes\id) (\ket{\Psi} \otimes \ket{\Omega}) = (\bra{\bar{\Psi}} \otimes\id) \ket{\Omega} = \ket{\Psi}.
\end{equation*}
We thus obtain that the state after Alice's measurement, if she obtains ouctcome $\alpha\beta$, is
\begin{equation*}
  \ket{\Phi_{\alpha\beta}} = \ket{\Omega_{\alpha\beta}}\otimes X^\alpha Z^\beta \ket{\Psi}.
\end{equation*}
Therefore, if Bob applies $Z^\beta X^\alpha$ on his part of the state, he will obtain the state $\ket{\Psi}$ independent of Alice's side.
